---
title: "Homework 3"
author: "Sonia Tandon"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: bookdown::html_document2
---

```{r, set-environment, include = FALSE}
#load packages 
lapply(c("haven", "ggplot2", "tidyverse", "VIM", "compareGroups", "Hmisc", "bookdown", "knitr"), library, character.only=TRUE)

#load data 
cmh <- read.csv("class-project-data.csv")

```
# View data 

```{r, view-data, echo = FALSE, message = FALSE, warning = FALSE}

glimpse(cmh)
```
There are `r nrow(cmh)` observations in the dataset, with `r ncol(cmh)` features measured on each. 

```{r, data-cleaning, include = FALSE, message = FALSE}
# remove redundant features, rename variables, convert sex to factor
cmh <- cmh %>% 
	select(-X, -FM_perc_11y, -FM_kg_11y, -FM_kg_11y_new, -FFM_perc_11y_new, -FFM_kg_11y_new) %>%
	rename(id = folio, sbp_11y = SISTO_11y_v2, dbp_11y = DIAST_11y_v2, weight_kg = WtPeso, fm_perc_11y = FM_perc_11y_new) %>% 
	mutate(sex = as.factor(sex)) 

levels(cmh$sex) <- c("Male", "Female")

#label data 
var.labels <- c(sex = "Sex",
				adiponec_11y = "Adiponectin",
				leptin_11y = "Leptin",
				il6_11y = "IL-6",
				insulin_11y = "Insulin",
				glucose_11y = "Glucose",
				cholester_11y = "Total Cholesterol",
				hdl_11y = "HDL",
				tgs_11y = "Triglycerides",
				ldl_11y = "LDL",
				sum_skin_11y = "Sum of Skinfolds",
				crp_11y = "C-reactive Protein",
				sbp_11y = "Systolic Blood Pressure, mmHg",
				dbp_11y = "Diastolic Blood Pressure, mmHg",
				weight_kg = "Weight, kg",
				fm_perc_11y = "Fat mass, %")

cmh <- Hmisc::upData(cmh, labels = var.labels)

```
# Descriptive Statistics

```{r, summary-statistics, echo = FALSE}
cmh_factors <- cmh %>% select(sex, adiponec_11y, leptin_11y, il6_11y, insulin_11y, glucose_11y, cholester_11y, hdl_11y, tgs_11y, ldl_11y, sum_skin_11y, crp_11y, sbp_11y, dbp_11y, weight_kg, fm_perc_11y)

table1 <- descrTable(select(cmh_factors, sex, adiponec_11y, leptin_11y, il6_11y, insulin_11y, glucose_11y, cholester_11y, hdl_11y, tgs_11y, ldl_11y, sum_skin_11y, crp_11y, sbp_11y, dbp_11y, weight_kg, fm_perc_11y))

export2md(table1)

```
As shown in Table \@ref(tab:summary-statistics), some features have missing data. 

# Descriptive statistics stratified by sex

```{r, stratified-table, echo = FALSE}
table2 <- compareGroups(sex ~ adiponec_11y + leptin_11y + il6_11y + insulin_11y + glucose_11y + cholester_11y + hdl_11y + tgs_11y + ldl_11y + sum_skin_11y + crp_11y + sbp_11y + dbp_11y + weight_kg + fm_perc_11y, data=cmh_factors)

table2_output <- createTable(table2)

export2md(table2_output)

```

# Patterns of missing data 

```{r, missing-data, echo = FALSE, warning = FALSE, message = FALSE}
aggr_plot <- aggr(cmh_factors, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(cmh), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
aggr_plot

```
Blood pressure (SBP, DBP) and inflammatory markers (IL-6, CRP) have the highest degree of missingness.








